A thesis draft

### 2018-11-21 res/sections/10-ProjectAnalysis.tex
Add new software descriptions and fix TODO

In conventional virtualization systems these resources are usually separated
from the host operating system, providing better isolation but worst
performance. The ETSI infrastructure, instead, tries with container
virtualization technology to push for a better NFV responsiveness\todo{Check
  term - I don't think it actually exists}, removing the additional operating
system (called ``guest'' OS) required by the usual hyper-visor-oriented 
approach.

@@

\item \textbf{VNFs} this are the core components that perform packets
  manipulation. This components are design to be lightweight services able to
  elaborate a great amount of packets per seconds. Manipulation can be of two
  typologies: stateful or stateless. In the first case, a state is maintained,
  and the VNF can perform heuristics about data coming through it (especially
  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
  case, no state is maintained, and every packet is treated without any
  particular assumption. Following the RFC 7665, the VNF are also called Service
  Function (SF)\footnote{From now on the two terminologies will be use in a
    interchangeable way.}. \todo{Add ref to chap about SFC?}

\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
  Orchestrator (NFVO) is responsible for the end-to-end management and
  orchestration of network services (NS), also called SFC. The MANO has
  different responsibilities, such as managing the
  creation/destruction/modification of the SFCs, that achieves interacting
  with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
Architectures: A Survey}. Finally, the MANO can integrate with 
Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
the following papers: https://www.tandfonline.com/doi/abs/10.1057/jors.1995.176 
https://ieeexplore.ieee.org/abstract/document/6963800 
https://www.computer.org/csdl/proceedings/ifcsta/2009/3930/01/3930a466-abs.html}

\end{itemize}

\section{SFC}

@@

\subsubsection{Openstack}
\label{chap:prjan:sec:openstack}
Created in ??\todo{Find out Openstack date of creation}, Openstack allows
on-premise cloud installations, using bare-metal resources to provide common
cloud services as object storage, virtual machine deploy, virtual networking.
Furthermore its modularity offers the possibility to add additional components,
even proprietary, to achieve a complete cloud solution.
\begin{figure}[t]
 \centering \includegraphics[scale=0.58]{openstack_logo}
 \caption{Openstack logo}
 \label{chap:prjan:img:openstack_logo}
\end{figure}

@@

As already introduced before, Kubernetes is an open source software framework
specialized in container management and orchestration. Developed by Google in
??\todo{Find out the actual date} and written in Go, it allows different
machines (called nodes from now on) to create an abstraction layer and to form a
cluster where is possible to run Docker containers without bothering about
hardware constraints. Nodes can have specific roles and specialization, all of
them managed by one or many master nodes, that actually manage the
orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
that it offers the possibility to change the virtualization software in use
(i.e. from Container-based technology to Virtual machine software).
\begin{figure}[h]
 \centering \includegraphics[scale=0.35]{kubernetes_logo}
 \caption{Kubernetes Logo}
 \label{chap:intro:img:k8s_logo}
\end{figure}

@@

In recent years, with the new hardware capabilities and the recent development
of in-kernel virtualization systems (such as Hyper-V) this technology begin to
be adopted. Virtualization allows to run different systems on the same machine,
making them completely isolated and then more resilient to failures. The back of
the medal, though, is that virtual machines require large amount of resources,
especially memory, because solutions like copy-and-write of in-memory pages are
not viable anymore (since the kernel gets duplicated too), leading to
duplication of loaded libraries and assets in the memory of the host. For large
deployment containing only simple services (e.g. a back-end application serving 
a
website or a database) this ends up in a waste of resources. \todo{Talk about
  lxc containers and how Docker solved magically the problem.} It's here that,
in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
on an already existing product: Linux Container (LXC). From this framework,
Docker built an entire ecosystem, consisting in a client/server model giving the
possibility for users to simply launch, scale and delete containers (locally or
remotely with Docker-machine), a repository system where images, a layer-based
``core'' where containers start running from when they are launched, can be
stored and tagged in a sort of version control system. Finally, another couple
of solutions, called Docker-compose, and Docker Swarm have the purpose to,
respectively, orchestrate containers in a single or a clustered system.
\begin{figure}[t]
 \centering \includegraphics[scale=0.7]{docker_logo}
 \caption{Official Docker logo}
 \label{chap:intro:img:docker_logo}
\end{figure}

@@

\subsubsection{Docker Swarm}
\label{chap:prjan:sec:tech:sub:other:sub:swarm}
Introduced in ??\todo{Find Docker Swarm date of birth}, Docker Swarm allows 
multiple Docker nodes to cluster together and be seen as one logical unit. This 
layer completely makes the underlying infrastructure transparent: data 
management, as network one, are totally managed by Docker. Recently, 
Docker-compose support has been introduced, making this one-node tool available 
for clustered systems. At the end of the day, on one hand Docker Swarm provides 
an easy way to set up a cluster system with all the tools configured 
out-of-the-box, without the necessity to set networking configurations or 
installing ad-hoc storage solutions. On the other hand, though, it doesn't have 
all the customization options that Kubernetes offers, thus making the product 
less flexible. This fundamental characteristic, although, makes the two 
solutions have different use cases leading to different market needs.

\subsubsection{OpenvSwitch} \todo{Talk about OpenvSwitch!!}
\subsubsection{Openbaton}
\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
Patronized by the Institute Fraunhofer, it's a open-source, customizable NFV 
MANO-compliant framework, that offers many configurations options, and since 
it's 
written in Java, there is the possibility to add components (plug-ins) 
dynamically. Openbaton can be installed as a normal computer program (they 
offer packages for the most common distros) but it can be deployed on Docker 
containers too. It's designed to work with cloud providers like Amazon or 
Google Compute engine, but it offers compatibility with on-premise cloud 
solutions like Openstack, where it deploys virtual machines using the API
the platform offers. It can store VNF configurations saved as TOSCA
YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
PoPs\todo{Write down PoP meaning}. It offers a VNF lifecycle management
out-of-the-box, with the possibility to customize it based on the user needs.
Its modular design allows developers to change parts of the codebase with custom
ones, making the product really flexible to different platforms. Performing this
operation requires a deep knowledge of the framework though.
\begin{figure}[h]
 \centering \includegraphics[scale=0.45]{openbaton_logo}
 \caption{Openbaton logo}
 \label{chap:prjan:img:openbaton_logo}
\end{figure}

>>>>

In conventional virtualization systems these resources are usually separated
from the host operating system, providing better isolation but worst
performance. The ETSI infrastructure, instead, tries with container
virtualization technology to push for a better NFV responsiveness, removing the
additional operating system (called ``guest'' OS) required by the usual
hyper-visor-oriented approach.

@@

\item \textbf{VNFs} this are the core components that perform packets
  manipulation. This components are design to be lightweight services able to
  elaborate a great amount of packets per seconds. Manipulation can be of two
  typologies: stateful or stateless. In the first case, a state is maintained,
  and the VNF can perform heuristics about data coming through it (especially
  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
  case, no state is maintained, and every packet is treated without any
  particular assumption. Following the RFC 7665, the VNFs are also called
  Service Function (SF)\footnote{From now on the two terminologies will be use
    in a interchangeable way.}. Additional infomation can be found in
  Section~\ref{chap:prjan:sfc}.

\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
  Orchestrator (NFVO) is responsible for the end-to-end management and
  orchestration of network services (NS), also called SFC. The MANO has
  different responsibilities, such as managing the
  creation/destruction/modification of the SFCs, that achieves interacting with
  the VIM~\cite{nguyen2017sdn}. Finally, the MANO can be integrated with
  Operational and Business Support Systems (OSS/BSS).
\end{itemize}

\section{SFC}
\label{chap:prjan:sfc}

@@

\subsubsection{Openstack}
\label{chap:prjan:sec:openstack}
Created in 2010 by a joint operation between Rackspace Hosting and
NASA~\cite{openstackWebsite}, Openstack allows on-premise cloud installations,
using bare-metal resources to provide common cloud services as object storage,
virtual machine deploy, virtual networking. Furthermore its modularity offers
the possibility to add additional components, even proprietary, to achieve a
complete cloud solution.
\begin{figure}[t]
 \centering \includegraphics[scale=0.58]{openstack_logo}
 \caption{Openstack logo}
 \label{chap:prjan:img:openstack_logo}
\end{figure}

@@

As already introduced before, Kubernetes is an open source software framework
specialized in container management and orchestration. Developed by Google in
2014 and written in Go~\cite{k8sGit}, it allows different machines (called nodes
from now on) to create an abstraction layer and to form a cluster where is
possible to run Docker containers without bothering about hardware constraints.
Nodes can have specific roles and specialization, all of them managed by one or
many master nodes, that actually manage the orchestration. On top of that,
Kubernetes is virtualization-agnostic, meaning that it offers the possibility to
change the virtualization software in use (i.e. from Container-based technology
to Virtual machine software).
\begin{figure}[h]
 \centering \includegraphics[scale=0.35]{kubernetes_logo}
 \caption{Kubernetes Logo}
 \label{chap:intro:img:k8s_logo}
\end{figure}

@@

In recent years, with the new hardware capabilities and the recent development
of in-kernel virtualization systems (such as Hyper-V) this technology begin to
be adopted. Virtualization allows to run different systems on the same machine,
making them completely isolated and then more resilient to failures. The back of
the medal, though, is that virtual machines require large amount of resources,
especially memory, because solutions like copy-and-write of in-memory pages are
not viable anymore (since the kernel gets duplicated too), leading to
duplication of loaded libraries and assets in the memory of the host. For large
deployment containing only simple services (e.g. a back-end application serving
a website or a database) this ends up in a waste of resources. \todo{Talk about
  lxc containers and how Docker solved magically the problem.} It is here that,
in 2013~\cite{dockerWebsite}, Docker was created, basing its solution on an
already existing product: Linux Container (LXC). From this framework, Docker
built an entire ecosystem, consisting in a client/server model giving the
possibility for users to simply launch, scale and delete containers (locally or
remotely with Docker-machine), a repository system where images, a layer-based
``core'' where containers start running from when they are launched, can be
stored and tagged in a sort of version control system. Finally, another couple
of solutions, called Docker-compose, and Docker Swarm have the purpose to,
respectively, orchestrate containers in a single or a clustered system.
\begin{figure}[t]
 \centering \includegraphics[scale=0.7]{docker_logo}
 \caption{Official Docker logo}
 \label{chap:intro:img:docker_logo}
\end{figure}

@@

\subsubsection{Docker Swarm}
\label{chap:prjan:sec:tech:sub:other:sub:swarm}
Introduced in 2014~\cite{swarmGit}, Docker Swarm allows multiple Docker nodes to
cluster together and be seen as one logical unit. This layer completely makes
the underlying infrastructure transparent: data management, as network one, are
totally managed by Docker. Recently, Docker-compose support has been introduced,
making this one-node tool available for clustered systems. At the end of the
day, on one hand Docker Swarm provides an easy way to set up a cluster system
with all the tools configured out-of-the-box, without the necessity to set
networking configurations or installing ad-hoc storage solutions. On the other
hand, though, it doesn't have all the customization options that Kubernetes
offers, thus making the product less flexible. This fundamental characteristic,
although, makes the two solutions have different use cases leading to different
market needs.

\subsubsection{OpenVSwitch}
Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an
open-source muti-layer virtual switch. It provides a switching stack for
hardware virtualization environments, implementing several interfaces and
protocols. It has also the possibility to run through Hypervisor virtualization
but it can run under Docker too~\cite{ovsDocker}. OpenVSwitch can route packets
from incoming connections following rules, that can be ``pushed'' in its routing
table from software defined as \textit{controllers}. Additionally, it supports the
OpenFlow communication protocol, in order to set previously cited routing
definitions.

\subsubsection{Floodlight}
\label{chap:prjan:sec:tech:sub:other:sub:floodlight}
Community-developed and open source, Floodlight is an OpenFlow Controller
written in Java at the end of 2011~\cite{floodlightGit}. It allows various
applications to run on top of it. Floodlight has the role of controlling and
inquiring an OpenFlow network, querying the applications integrated with the
purpose of solving different user needs over it. When instantiated, OpenSwitch
routers have to register to the controller in order to gain routing rules.

\subsubsection{Openbaton}
\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
Patronized by the Institute Fraunhofer, it is an open-source, customizable NFV
MANO-compliant framework, that offers many configurations options, and since
it's written in Java, there is the possibility to add components (plug-ins)
dynamically. Openbaton can be installed as a normal computer program (they offer
packages for the most common distros) but it can be deployed on Docker
containers too. It's designed to work with cloud providers like Amazon or Google
Compute engine, but it offers compatibility with on-premise cloud solutions like
Openstack, where it deploys virtual machines using the API the platform offers.
It can store VNF configurations saved as Topology and Orchestration
Specification for Cloud Applications (TOSCA) Yet Another Markup Language (YAML)
and it can handle multiple Point of Presence (PoP). It offers a VNF lifecycle
management out-of-the-box, with the possibility to customize it based on the
user needs. Its modular design allows developers to change parts of the codebase
with custom ones, making the product really flexible to different platforms.
Performing this operation requires a deep knowledge of the framework though.
\begin{figure}[h]
 \centering \includegraphics[scale=0.45]{openbaton_logo}
 \caption{Openbaton logo}
 \label{chap:prjan:img:openbaton_logo}
\end{figure}
